---
title: "6030_project"
author: "Claire Sullivan, Terrance Luangrath, Katie Dunning, Samantha Asef"
date: "2025-06-13"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_chunk$set(cache=TRUE, autodep=TRUE)
knitr::opts_chunk$set(fig.align="center", fig.pos="tbh")
library(tidyverse)
library(tidymodels)
library(discrim)
library(patchwork)
library(glmnet)
library(future)
library(dplyr)
library(knitr)
library(kknn)
```

```{r setup parallel, include=FALSE, cache=FALSE}
library(future)
plan(multisession, workers = parallel::detectCores(logical = FALSE) - 2)
```

#PreProcessing

```{r}
# | message: false
# | warning: false
HaitiTrain<-read_csv("HaitiPixels.csv")
```
```{r}
HaitiTrain<-HaitiTrain%>%
  mutate(Class=ifelse(Class=="Blue Tarp", 1, 0))%>%
  mutate(Class=factor(Class, levels=c(1,0), labels=c("Blue_Tarp", "Other")))
```

```{r, message=FALSE}
name_column=c("ID","X","Y","Map_X","Map_y","Lat","Lon","Red","Green","Blue")

OT1<-read_table("HoldOutData/orthovnir067_ROI_NOT_Blue_Tarps.txt", skip=8, col_names = name_column)

OT2<-read_table("HoldOutData/orthovnir069_ROI_NOT_Blue_Tarps.txt", skip=8, col_names = name_column)

OT3<-read_table("HoldOutData/orthovnir078_ROI_NON_Blue_Tarps.txt", skip=8, col_names = name_column)

BT1<-read_table("HoldOutData/orthovnir067_ROI_Blue_Tarps.txt", skip=8, col_names = name_column)

BT2<-read_table("HoldOutData/orthovnir069_ROI_Blue_Tarps.txt", skip=8, col_names = name_column)

BT3<-read_table("HoldOutData/orthovnir078_ROI_Blue_Tarps.txt", skip=8, col_names = name_column)
```

```{r}
Blue_Tarp<-bind_rows(BT1,BT2,BT3)%>%
  mutate(class=1)

Other<-bind_rows(OT1,OT2,OT3)%>%
  mutate(class=0)

Testdat <-bind_rows(Blue_Tarp, Other)%>%
  mutate(class=factor(class, levels=c(1,0), labels=c("Blue_Tarp", "Other")))

Testdat <- Testdat %>% rename(Class = class)
```


```{r density, fig.height=2.5, fig.width=8}

g1<-ggplot(HaitiTrain)+

  geom_density(aes(x=Red, color="red"), alpha=0.7)+

  geom_density(aes(x=Green, color="green"), alpha=0.7)+

  geom_density(aes(x=Blue, color="blue"), alpha=0.7)+

  scale_color_manual(values=c("red"="red","green"="green","blue"="blue"),

  limits=c("red","green","blue"))+

  labs(title="Known RGB Haiti Train",x="RGB Values",y="Density")
 
g2<-ggplot(Testdat)+

  geom_density(aes(x=Red, color="red"), alpha=0.7)+

  geom_density(aes(x=Green, color="green"), alpha=0.7)+

  geom_density(aes(x=Blue, color="blue"), alpha=0.7)+

  scale_color_manual(values=c("red"="red","green"="green","blue"="blue"),

  limits=c("red","green","blue"))+

  labs(title="Suspected RGB Haiti Test",x="RGB Values",y="Density")
 
(g1+g2)

```
 
# Explotatory Data Analysis 

```{r}
#check for NA values
sum(is.na(HaitiTrain))
```

```{r}
#check for NA values
sum(is.na(Testdat))
```


```{r}
HaitiTrain %>%
  count(Class)
```
```{r}
Testdat %>%
  count(Class)
```

```{r}
g1 <- Testdat %>% 
  ggplot(aes(x=Class, fill = Class)) + 
  scale_y_log10() + 
  scale_fill_manual(values = c("lightblue", "grey"), guide = "none") + 
  geom_bar() + 
  theme_bw() + 
  labs(x= "Class", y="Count", title = "Holdout Data")
g2 <- HaitiTrain %>% 
  ggplot(aes(x=Class, fill = Class)) + 
  scale_y_log10() + 
  scale_fill_manual(values = c("lightblue", "grey"), guide = "none") + 
  geom_bar() + 
  theme_bw() + 
  labs(x= "Class", y="Count", title = "Test Data")

g1 + g2
```



```{r}
#distribution of RGB for Blue Tarp and Other 
library(ggplot2)
library(dplyr)
library(tidyr)

# one RGB column
HaitiLong <- HaitiTrain %>%
  pivot_longer(cols = c(Red, Green, Blue), names_to = "Colors", values_to = "Value")
TestLong <- Testdat %>%
  pivot_longer(cols = c(Red, Green, Blue), names_to = "Colors", values_to = "Value")

g3 <- ggplot(HaitiLong, aes(x = Value, color = Colors)) +
  geom_density() +
  facet_wrap(~Class) +
  scale_color_manual(values = c("Red" = "red", "Green" = "green", "Blue" = "blue")) +
  labs(color = "Color", x = "Values", y = "Density", title = "Training Data") +
  theme_bw()

g4 <- ggplot(TestLong, aes(x = Value, color = Colors)) +
  geom_density() +
  facet_wrap(~Class) +
  scale_color_manual(values = c("Red" = "red", "Green" = "green", "Blue" = "blue")) +
  labs(color = "Color", x = "Values", y = "Density", title = "Holdout Data") +
  theme_bw()

(g3 + g4) + plot_layout(guides = "collect")
```

```{r}
HaitiRatio <- HaitiTrain %>% 
  mutate(BlueGreen = (Blue/Green)) %>% 
  mutate(BlueRed = (Blue/Red)) %>% 
  select(-Blue, -Red, -Green)
```

```{r}
TestRatio <- Testdat %>% 
  mutate(BlueGreen = (Blue/Green)) %>% 
  mutate(BlueRed = (Blue/Red)) %>% 
  select(-Blue, -Red, -Green)
```
```{r}
HaitiRatio2 <- HaitiRatio %>% 
  pivot_longer(cols = c(BlueGreen, BlueRed), names_to = "Color_Ratios", values_to = "Value")
```
```{r}
TestRatio2 <- TestRatio %>% 
  pivot_longer(cols = c(BlueGreen, BlueRed), names_to = "Color_Ratios", values_to = "Value")
```


```{r}
ggplot(HaitiRatio2, aes(x = Class, y = Value)) + 
  geom_boxplot()+ 
  facet_wrap(~Color_Ratios) + 
  theme_bw()
```


```{r}
g5 <- ggplot(HaitiRatio, aes(x = BlueGreen, y = BlueRed, color = Class)) + 
  geom_point(alpha = 0.1)+ 
  scale_color_manual(values = c("blue", "darkgrey"))+
  theme_bw()
g6 <- ggplot(TestRatio, aes(x = BlueGreen, y = BlueRed, color = Class)) + 
  geom_point(alpha = 0.1)+ 
  scale_color_manual(values = c("blue", "darkgrey"))+
  theme_bw()

(g5 + g6) + plot_layout(guides = "collect")
```
```{r}
g5 <- ggplot(HaitiRatio, aes(x = BlueGreen, y = BlueRed, color = Class)) + 
  geom_density_2d(bins=20) +
  scale_color_manual(values = c("lightblue", "darkgrey"))+
  theme_bw() + 
  labs(x = "BlueGreen", y = "BlueRed", title = "Training Data")
g6 <- ggplot(TestRatio, aes(x = BlueGreen, y = BlueRed, color = Class)) + 
  geom_density_2d(bins=20) +
  scale_color_manual(values = c("lightblue", "darkgrey"))+
  theme_bw()+ 
  labs(x = "BlueGreen", y = "BlueRed", title = "Holdout Data")
g7 <- ggplot(HaitiRatio, aes(x = BlueGreen, y = BlueRed, color = Class)) + 
  geom_point(alpha = 0.1)+ 
  scale_color_manual(values = c("lightblue", "darkgrey"), guide = "none")+
  theme_bw()+ 
  labs(x = "BlueGreen", y = "BlueRed", title = "Training Data")
g8 <- ggplot(TestRatio, aes(x = BlueGreen, y = BlueRed, color = Class)) + 
  geom_point(alpha = 0.1)+ 
  scale_color_manual(values = c("lightblue", "darkgrey"), guide = "none")+
  theme_bw()+ 
  labs(x = "BlueGreen", y = "BlueRed", title = "Holdout Data")

((g5 + g6) + plot_layout(guides = "collect")) / ((g7+ g8) + plot_layout(guides = "collect")) 
```


# Tuning Models: 

### Recipe, Model Specs
```{r}
sapply(HaitiTrain, class)
```
```{r}
formula <- Class ~ . 
rec <- recipe(formula, data=HaitiTrain) %>%
    step_normalize(all_numeric_predictors()) %>% 
    step_dummy(all_nominal_predictors())
```
```{r}
#LDA 
lda_spec <- discrim_linear(mode = "classification", engine = "MASS")

#QDA 
qda_spec <- discrim_quad(mode = "classification", engine = "MASS")

#KNN 
knn_spec <- nearest_neighbor(engine="kknn", mode="classification",neighbors=tune())
#LR
logreg_spec <- logistic_reg(engine="glmnet", mode="classification", penalty=tune(), mixture=tune())
#RF
rf_spec <- rand_forest(
  mode = "classification",
  mtry = tune(),
  trees = 1000,
  min_n = tune()
  )
#SVM
poly_spec <- svm_poly(mode = "classification",
                      cost = tune(),
                      scale_factor = tune(), 
                      degree = tune()) %>% 
  set_engine("kernlab")

```

### Workflow
```{r}
#LDA 
lda_workflow <- workflow() %>%
  add_recipe(rec) %>%
  add_model(lda_spec)

#QDA 
qda_workflow <- workflow() %>%
  add_recipe(rec) %>%
  add_model(qda_spec)

#KNN 
knn_wf <- workflow() %>%
    add_recipe(rec) %>%
    add_model(knn_spec)
#LR
logreg_wf <- workflow() %>%
    add_recipe(rec) %>%
    add_model(logreg_spec)
#RF
rf_wf <- workflow() %>% 
  add_recipe(rec) %>% 
  add_model(rf_spec %>% 
              set_engine("ranger", importance="impurity"))
#SVM
svm_wf <- workflow() %>% 
  add_recipe(rec) %>% 
  add_model(poly_spec)
```

### Defining Parameters 

```{r}
#LDA 
# No parameters for LDA

#QDA 
# No parameters for QDA

#KNN 
knn_params <- extract_parameter_set_dials(knn_wf) %>%
    update(neighbors=neighbors(c(2, 40)))
#LR
logreg_params <- extract_parameter_set_dials(logreg_wf) %>%
    update(
        penalty=penalty(c(-3, 1)),
        mixture=mixture(c(0, 1)))
#RF
rf_params <- extract_parameter_set_dials(rf_wf) %>%
  update(
    mtry=mtry(c(1, 3)),
    min_n=min_n(c(2, 40))
  )
#SVM
svm_params <- extract_parameter_set_dials(svm_wf) %>%
  update(
    cost=cost(c(-10, 5)),
    degree=degree(c(1, 3)), 
    scale_factor=scale_factor(c(-10, -1))
  )
```

### CV Folds, Resamples, Controls 

```{r}
set.seed(6030)
resamples <- vfold_cv(HaitiTrain, v=10, strata=Class)
mycontrols = control_resamples(save_pred = TRUE)
#saves OOF
mymetrics = metric_set(roc_auc, f_meas)
```

### Tuning Hyperparameters
```{r, message=FALSE, warning=FALSE}
#library(doParallel)
#registerDoParallel(cores = parallel::detectCores() - 1)
set.seed(6030)
#KNN
nn_results <- tune_grid(knn_wf,
    resamples=resamples,
    control=mycontrols,
    metrics = mymetrics,
    grid=grid_latin_hypercube(knn_params, size = 20))
#LR
logreg_results <- tune_grid(logreg_wf,
    resamples=resamples,
    control=mycontrols,
    metrics = mymetrics,
    grid=grid_latin_hypercube(logreg_params, size = 20))

#RF
rf_results <- tune_grid(
  rf_wf,
  resamples=resamples,
  control=mycontrols,
  metrics=mymetrics,
  grid=grid_latin_hypercube(rf_params, size = 20)
)

#SVM
svm_results <- tune_grid(
  svm_wf,
  resamples=resamples,
  control=mycontrols,
  metrics=mymetrics,
  grid=grid_latin_hypercube(svm_params, size = 20)
)
```

### Evaluating Initial Parameter Range

```{r}
#KNN
knn_best_params<- show_best(nn_results, metric="roc_auc", n=1)
```
```{r}
knn_best_params
```
```{r}
autoplot(nn_results)
```
```{r}
#LR
logreg_best_params <- show_best(logreg_results, metric="roc_auc", n=1)
```
```{r}
logreg_best_params
```
```{r}
autoplot(logreg_results)
```
```{r}
#RF 
rf_best_params <- show_best(rf_results, metric="roc_auc", n=1)
rf_best_params
```
```{r}
autoplot(rf_results) +
  labs(title="Random Forest Tuning Results")
```
```{r}
#SVM
svm_best_params <- show_best(svm_results, metric="roc_auc", n=1)
svm_best_params
```
```{r}
autoplot(svm_results) + 
  labs(title="SVM-Poly Tuning Results")
```



### RE- Defining Parameters 

```{r}
#KNN
knn_params <- extract_parameter_set_dials(knn_wf) %>%
    update(neighbors=neighbors(c(2, 20)))
#LR
logreg_params <- extract_parameter_set_dials(logreg_wf) %>%
    update(
        penalty=penalty(c(-5, 1)),
        # penalty is on a log 10 scale 
        mixture=mixture(c(0, 1)))
        # based on a mixture of L1 and L2 
```

### Tuning Hyperparameters - with updated Parameters and CV resamples (LDA/QDA)

```{r, message=FALSE, warning=FALSE}
#library(doParallel)
#registerDoParallel(cores = parallel::detectCores() - 1)
set.seed(6030)
#LDA
lda_cv <- fit_resamples(lda_workflow, resamples, 
                        metrics=mymetrics, 
                        control=mycontrols)
#QDA
qda_cv <- fit_resamples(qda_workflow, 
                        resamples, 
                        metrics=mymetrics, 
                        control=mycontrols)
#KNN
nn_results <- tune_grid(knn_wf,
    resamples=resamples,
    control=mycontrols,
    metrics = mymetrics,
    grid=grid_latin_hypercube(knn_params, size = 20))
#LR
logreg_results <- tune_grid(logreg_wf,
    resamples=resamples,
    control=mycontrols,
    metrics = mymetrics,
    grid=grid_latin_hypercube(logreg_params, size = 20))
```

### Evaluating Updated Parameter Range
```{r}
#KNN
knn_best_params<- show_best(nn_results, metric="roc_auc", n=1)
```
```{r}
knn_best_params
```
```{r}
autoplot(nn_results)
```
```{r}
#LR
logreg_best_params <- show_best(logreg_results, metric="roc_auc", n=1)
```
```{r}
autoplot(logreg_results)
```
```{r}
logreg_best_params
```

### Collect predictions for out-of-fold predictions

```{r}
#LDA and QDA
lda_preds<-collect_predictions(lda_cv)
qda_preds<-collect_predictions(qda_cv)
```

```{r}
#KNN 
KNN_preds <- collect_predictions(nn_results) %>% 
  filter(neighbors == knn_best_params$neighbors) 
#LR
LR_preds <- collect_predictions(logreg_results) %>% 
  filter(penalty == logreg_best_params$penalty, mixture == logreg_best_params$mixture) 
#RF
rf_preds <- collect_predictions(rf_results) %>% 
  filter(mtry == rf_best_params$mtry, min_n == rf_best_params$min_n)
#SVM
svm_preds <-collect_predictions(svm_results) %>% 
  filter(cost == svm_best_params$cost, 
         degree == svm_best_params$degree,
         scale_factor == svm_best_params$scale_factor)

```

### CV ROC Curve

```{r}
roc_cv_data<-function(model_pred, model_name) {
  roc_data<-model_pred%>%
    roc_curve(truth=Class, .pred_Blue_Tarp, event_level="first")%>%
    mutate(Model=model_name)
  return(roc_data)
}
```

```{r}
roc_dat<-bind_rows(
  roc_cv_data(lda_preds, "LDA Model"),
  roc_cv_data(qda_preds, "QDA Model"),
  roc_cv_data(KNN_preds, "KNN Model"),
  roc_cv_data(LR_preds, "Pen Logistic Mod"),
  roc_cv_data(rf_preds, "Random Forest Mod"),
  roc_cv_data(svm_preds, "SVM Polynomial Mod")
)

roc_cv_g<-ggplot(roc_dat, aes(x=1 - specificity, y=sensitivity, color=Model))+
                   geom_line()+
                   theme(aspect.ratio = 1)

roc_cv_g_zoom <- roc_cv_g+
  coord_cartesian(xlim=c(0,0.15),
                  ylim=c(0.85,1))+
                  guides(color="none")

roc_cv_g+roc_cv_g_zoom
```

### Thresholding on OOF predictions 

```{r function 1}

# Threshold graphs function

threshold_graph<-function(OOF,model_name){
  performance<-probably::threshold_perf(OOF, Class, .pred_Blue_Tarp,
    thresholds=seq(0.001,0.5,0.001), event_level="first",
    metrics=metric_set(f_meas,accuracy,j_index,sensitivity))
  max_metrics<-performance%>%
    drop_na()%>%
    group_by(.metric)%>%
    filter(.estimate==max(.estimate))
 g<-ggplot(performance,aes(x=.threshold,y=.estimate,color=.metric))+
   geom_line()+
   geom_point(data=max_metrics, color="black")+
   labs(title=model_name,x="Threshold",y="Metricvalue")
   #coord_cartesian(ylim=c(0,0.825))
 thresholds<-max_metrics%>%
   dplyr::select(.metric,.threshold)%>%
   deframe()
 return(list(graph=g,thresholds=thresholds))
 }
```

```{r function 2}

# Confusion Matrix Function

 visualize_conf_mat<-function(OOF,thresholds,metric){
   threshold<-thresholds[metric]
   cm<-OOF%>%
     mutate(
       .pred_class=probably::make_two_class_pred(.pred_Blue_Tarp,c("Blue_Tarp","Other"),
                                                 threshold=threshold),
       )%>%
     conf_mat(truth=Class,estimate=.pred_class)
 autoplot(cm,type="heatmap")+
 labs(title=sprintf("Threshold%.2f(%s)",threshold,metric))
 }
```

```{r function 3}

# Putting graphs and model together

 overview_model<-function(OOF,model_name){
 tg<-threshold_graph(OOF,model_name)
 g1<-visualize_conf_mat(OOF,tg$thresholds,"accuracy")
 g2<-visualize_conf_mat(OOF,tg$thresholds,"f_meas")
 g3<-visualize_conf_mat(OOF,tg$thresholds,"j_index")
 g4<-visualize_conf_mat(OOF,tg$thresholds,"sensitivity")
 tg$graph+(g1/g2/g3/g4)
 }
```

```{r}

# Making visualization for KNN and Logreg with all ideal thresholds.
#LDA 
gldathresh<-overview_model(lda_preds, "LDA Model")

#QDA 
gqdathresh<-overview_model(qda_preds, "QDA Model")

#KNN 
gknnthresh<-overview_model(KNN_preds, "KNN Model")

#LR
glogithresh<-overview_model(LR_preds, "Pen Logistic Model")

#RF
grfthresh <- overview_model(rf_preds, "Random Forest")

#SVM
gsvmthresh <- overview_model(svm_preds, "SVM Poly")
```
```{r, fig.width=8, fig.height=14}
(gldathresh/gqdathresh/gknnthresh)
```
```{r, fig.width=8, fig.height=14}
(glogithresh/grfthresh/gsvmthresh)
```

### Combined threshold graph

```{r}
plot_optimal_f1 <- function(OOF, model_name){
  probably::threshold_perf(
    .data=OOF,
    truth=Class,
    estimate=.pred_Blue_Tarp,
    thresholds=seq(0.001, 0.5, 0.001),
    metrics=metric_set(f_meas)
  ) %>% 
    filter(.metric=="f_meas") %>%
    transmute(
      threshold=.threshold,
      f1=.estimate,
      model=model_name
    )
}

all_f1 <- bind_rows(
  plot_optimal_f1(lda_preds, "LDA"),
  plot_optimal_f1(qda_preds, "QDA"),
  plot_optimal_f1(KNN_preds,  "KNN"),
  plot_optimal_f1(LR_preds, "Penalized LR"),
  plot_optimal_f1(rf_preds, "Random Forest"),
  plot_optimal_f1(svm_preds, "SVM Poly")
)

ggplot(all_f1, aes(x = threshold, y = f1, color = model)) +
  geom_line() +
  geom_point(
    data=all_f1 %>%
      group_by(model) %>%  
      filter(f1 == max(f1)),
    aes(x=threshold, y=f1),
    colour="black",
    size=3) +
  labs(
    title="Optimal F1 vs Threshold by Model",
    x="Threshold",
    y="F1 Score"
  )
```

### Extract F-measure thresholds

```{r}
# Make metric object
ldathreshtest <- probably::threshold_perf(
  lda_preds,
  truth = Class,
  estimate = .pred_Blue_Tarp,
  thresholds = seq(0.001, 0.5, 0.001),
  metrics = metric_set(accuracy, f_meas, j_index, sensitivity))

#QDA 
qdathreshtest <- probably::threshold_perf(
  qda_preds,
  truth = Class,
  estimate = .pred_Blue_Tarp,
  thresholds = seq(0.001, 0.5, 0.001),
  metrics = metric_set(accuracy, f_meas, j_index, sensitivity))

#KNN 
knnthreshtest <- probably::threshold_perf(
  KNN_preds,
  truth = Class,
  estimate = .pred_Blue_Tarp,
  thresholds = seq(0.001, 0.5, 0.001),
  metrics = metric_set(accuracy, f_meas, j_index, sensitivity))

#LR
penlogithreshtest <- probably::threshold_perf(
  LR_preds,
  truth = Class,
  estimate = .pred_Blue_Tarp,
  thresholds = seq(0.001, 0.5, 0.001),
  metrics = metric_set(accuracy, f_meas, j_index, sensitivity))

#RF
rfthreshtest <- probably::threshold_perf(
  rf_preds,
  truth = Class,
  estimate = .pred_Blue_Tarp,
  thresholds = seq(0.001,0.5,0.001),
  metrics = metric_set(accuracy, f_meas, j_index, sensitivity))

#SVM
svmthreshtest <- probably::threshold_perf(
  svm_preds,
  truth = Class,
  estimate = .pred_Blue_Tarp,
  thresholds = seq(0.001,0.5,0.001),
  metrics = metric_set(accuracy, f_meas, j_index, sensitivity))
```

```{r}
# Pull F-measure threshold  

#LDA 
lda_opt<-ldathreshtest%>%
  filter(.metric=="f_meas")%>%
  arrange(-.estimate)%>%
  first()

lda_best_thresh<-lda_opt$.threshold

#QDA 
qda_opt<-qdathreshtest%>%
  filter(.metric=="f_meas")%>%
  arrange(-.estimate)%>%
  first()

qda_best_thresh<-qda_opt$.threshold

#KNN
knn_opt<-knnthreshtest%>%
  filter(.metric=="f_meas")%>%
  arrange(-.estimate)%>%
  first()

knn_best_thresh<-knn_opt$.threshold

#LR 
penlogi_opt<-penlogithreshtest%>%
  filter(.metric=="f_meas")%>%
  arrange(-.estimate)%>%
  first()

penlogi_best_thresh<-penlogi_opt$.threshold

#RF
rf_opt <- rfthreshtest %>%
  filter(.metric=="f_meas") %>%
  arrange(-.estimate) %>%
  first()

rf_best_thresh<-rf_opt$.threshold

#SVM
svm_opt <- svmthreshtest %>%
  filter(.metric=="f_meas") %>%
  arrange(-.estimate) %>%
  first()

svm_best_thresh<-svm_opt$.threshold

```

### Finalize WF & Fit train data

```{r}
#LDA 
lda_final_model <- fit(lda_workflow, data = HaitiTrain)

#QDA 
qda_final_model <- fit(qda_workflow, data = HaitiTrain)

#KNN
knn_final_model<- finalize_workflow(knn_wf, select_best(nn_results, metric = "roc_auc")) %>%
  fit(data = HaitiTrain)
#LR
logreg_final_model<- finalize_workflow(logreg_wf, select_best(logreg_results, metric = "roc_auc")) %>%
  fit(data = HaitiTrain)
#RF
rf_final_model <- finalize_workflow(rf_wf, select_best(rf_results, metric="roc_auc")) %>% 
  fit(data=HaitiTrain)
#SVM
svm_final_model <- finalize_workflow(svm_wf, select_best(svm_results, metric="roc_auc")) %>%
  fit(data=HaitiTrain)
```

### CV performance with resamples

```{r}
knn_resamp_only<-finalize_workflow(knn_wf, select_best(nn_results, metric = "roc_auc"))
logreg_resamp_only<-finalize_workflow(logreg_wf, select_best(logreg_results, metric = "roc_auc"))
rf_resamp_only<-finalize_workflow(rf_wf, select_best(rf_results, metric="roc_auc"))
svm_resamp_only<-finalize_workflow(svm_wf, select_best(svm_results, metric="roc_auc"))

knn_resamp<-fit_resamples(knn_resamp_only, resamples, metrics=mymetrics, control=mycontrols)
logreg_resamp<-fit_resamples(logreg_resamp_only, resamples, metrics=mymetrics, control=mycontrols)
rf_resamp<-fit_resamples(rf_resamp_only, resamples, metrics=mymetrics, control=mycontrols)
svm_resamp<-fit_resamples(svm_resamp_only, resamples, metrics=mymetrics, control=mycontrols)

bind_rows(
        collect_metrics(lda_cv) %>%
        mutate(model="LDA"),
        collect_metrics(qda_cv) %>%
        mutate(model="QDA"),
        collect_metrics(knn_resamp)%>%
        mutate(model="KNN"),
        collect_metrics(logreg_resamp)%>%
        mutate(model="Pen Logistic"),
        collect_metrics(rf_resamp)%>%
        mutate(model="Random Forest"),
        collect_metrics(svm_resamp)%>%
        mutate(model="SVM Polynomial")
)%>%
  select(model, .metric, mean)%>%
  pivot_wider(names_from = .metric, values_from = mean)%>%
  knitr::kable(caption="Cross-Validiation Performance All Models")
```


### Predicted probabilities with test data 

### Apply optimized threshold to the test preds to get a classification for test data 

```{r thresh forms}
predict_at_threshold <- function(model, data, threshold) {
    return(
        model %>%
            augment(data) %>%
            mutate(.pred_class = probably::make_two_class_pred(.pred_Blue_Tarp,
                    c("Blue_Tarp","Other"), threshold=threshold)))}

calculate_metrics_at_threshold <- function(model, train, test, model_name, threshold, met_set) {
    bind_rows(
      bind_cols(
            model=model_name, dataset="train", threshold=threshold,
            met_set(predict_at_threshold(model, train, threshold), 
                truth=Class, estimate=.pred_class, event_level="first"),
        ),
    bind_cols(model=model_name,dataset="train",threshold=threshold,
              roc_auc(model %>%augment(train),Class,.pred_Blue_Tarp,event_level="first"),
        ),
    bind_cols(
            model=model_name, dataset="test", threshold=threshold,
            met_set(predict_at_threshold(model, test, threshold), 
                truth=Class, estimate=.pred_class, event_level="first"),
        ),
    bind_cols(
      model=model_name,dataset="test",threshold=threshold,
      roc_auc(model %>%augment(test),Class,.pred_Blue_Tarp,event_level="first"),))}

 metrics_table<-function(model_metrics,caption){
   model_metrics %>%
     pivot_wider(names_from=.metric,values_from=.estimate)%>%
     select(-.estimator)%>%
     knitr::kable(caption=caption,digits=3,longtable=FALSE) %>%
     kableExtra::kable_styling(full_width=FALSE) }
```

```{r}
met_set<-metric_set(accuracy,kap,j_index,f_meas,sensitivity)

# ADD LDA, QDA, RANDOM FOREST, SVM
metrics_at_threshold<-bind_rows(  
  calculate_metrics_at_threshold(lda_final_model, HaitiTrain, Testdat, "LDA Model", lda_best_thresh,met_set),
  calculate_metrics_at_threshold(qda_final_model, HaitiTrain, Testdat, "QDA Model", qda_best_thresh,met_set),
  calculate_metrics_at_threshold(knn_final_model, HaitiTrain, Testdat, "KNN Model", knn_best_thresh,met_set),
calculate_metrics_at_threshold(logreg_final_model, HaitiTrain, Testdat, "Penalized Logreg Model", penlogi_best_thresh,met_set),
  calculate_metrics_at_threshold(rf_final_model, HaitiTrain, Testdat, "Random Forest Mod",
rf_best_thresh,met_set),
  calculate_metrics_at_threshold(svm_final_model, HaitiTrain, Testdat, "SVM Poly Model",
svm_best_thresh,met_set)
)

metrics_table(metrics_at_threshold, "Performance Metrics for All Models at Thresholds")
```

### Graph for the difference in train and test

```{r}
graph_fscore<-metrics_at_threshold%>%
  pivot_wider(names_from=.metric,values_from=.estimate)%>%
  select(model,dataset,f_meas)%>%
  pivot_wider(names_from=dataset, values_from=f_meas)%>%
  mutate(difference=train-test)%>%
  mutate(model=factor(model, levels=c("Random Forest Mod", "LDA Model",
                            "KNN Model", "SVM Poly Model",
                            "Penalized Logreg Model","QDA Model")))
```

```{r}
ggplot(graph_fscore, aes(x=model, y=difference, fill=model))+
  geom_bar(stat="identity")+
  labs(x="Model",y="F-Measure Difference",
       title="Training/Testing Differences in  F-Measure")+
  scale_fill_manual(values=c("Random Forest Mod"="firebrick",
                             "LDA Model"="firebrick",
                             "KNN Model"="firebrick",
                             "SVM Poly Model"="firebrick",
                             "Penalized Logreg Model"="cadetblue",
                             "QDA Model"="cadetblue"))+
  geom_smooth(aes(group=1),method="loess", se=FALSE,
              color="black")+
  theme(axis.text.x=element_text(angle=45, hjust=1))
  
```

### Heatmap Graph for Test data

```{r}
metrics_wide <- metrics_at_threshold %>%
  filter(dataset=="test") %>% 
  select(model, dataset, .metric, .estimate) %>% 
  pivot_wider(
    names_from=.metric,
    values_from=.estimate
  )

perf_long <- metrics_wide %>%
  pivot_longer(
    cols = -c(model, dataset),
    names_to  = "metric",
    values_to = "value"
  )

ggplot(perf_long, aes(x = metric, y = model, fill = value)) +
  geom_tile(color="white", size=0.2) +
  scale_fill_viridis_c(option = "A", name = NULL, limits = c(0, 1)) +
  labs(
    title = "Test Set Performance by Metric & Dataset",
    x=NULL,
    y=NULL
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid       = element_blank())
```

### Precision recall curve for test data
```{r}
bind_rows(
  augment(lda_final_model,Testdat)%>%
    mutate(model="LDA Model"),
  augment(qda_final_model,Testdat)%>%
    mutate(model="QDA Model"),
  augment(knn_final_model,Testdat)%>%
    mutate(model="KNN Model"),
  augment(logreg_final_model,Testdat)%>%
    mutate(model="Pen Logistic Model"),
  augment(rf_final_model,Testdat)%>%
    mutate(model="Random Forest Mod"),
  augment(svm_final_model,Testdat)%>%
    mutate(model="SVM Poly Model")
)%>%
  group_by(model)%>%
  pr_curve(truth=Class, .pred_Blue_Tarp, event_level="first")%>%
  autoplot()+
  geom_abline(intercept = 1, slope = -1, linetype="dashed")+
  labs(x="recall",y="precision",
       title="Precision-Recall Curve for Test Data")
```

### Scatter Graph for Final Data

```{r}
testmets1 <- metric_set(f_meas, kap, j_index)

testmets2 <- function(model, model_name) {
  bind_rows(
    testmets1(augment(model, HaitiTrain), truth = Class, estimate = .pred_class) %>% 
    mutate(model = model_name, dataset = "Train"),
    testmets1(augment(model, Testdat), truth = Class, estimate = .pred_class) %>% 
    mutate(model = model_name, dataset = "Test"),
  )
}
```

```{r}
ldascatter <- testmets2(lda_final_model, "LDA")
qdascatter <- testmets2(qda_final_model, "QDA")
knnscatter <- testmets2(knn_final_model, "KNN")
lrscatter<- testmets2(logreg_final_model, "Log Reg")
rfscatter<- testmets2(rf_final_model, "Random Forest")
svmscatter <- testmets2(svm_final_model, "SVM")
 
allmets <- bind_rows(ldascatter, qdascatter, knnscatter, lrscatter, rfscatter, svmscatter)

allmets<- allmets %>% 
  pivot_wider(names_from=.metric,values_from=.estimate)%>%
  select(-.estimator)

```

```{r, fig.width=8, fig.height=14}
gs1 <- ggplot(allmets, aes(x=f_meas, y=model, color=dataset)) +
    geom_point() + 
  theme_bw()

gs2 <- ggplot(allmets, aes(x=kap, y=model, color=dataset)) +
    geom_point() + 
  theme_bw()

gs3 <- ggplot(allmets, aes(x=j_index, y=model, color=dataset)) +
    geom_point() + 
  theme_bw()

gs1 / gs2 / gs3 
```

### Computer performance on the test data 

Confusion matrix, etc. here 

```{r}
conf_mat_lda<-predict_at_threshold(lda_final_model, Testdat, lda_best_thresh)%>%
  conf_mat(truth=Class, estimate=.pred_class)
conf_mat_qda<-predict_at_threshold(qda_final_model, Testdat, qda_best_thresh)%>%
  conf_mat(truth=Class, estimate=.pred_class)
conf_mat_knn<-predict_at_threshold(knn_final_model, Testdat, knn_best_thresh)%>%
  conf_mat(truth=Class, estimate=.pred_class)
conf_mat_logreg<-predict_at_threshold(logreg_final_model, Testdat, penlogi_best_thresh)%>%
  conf_mat(truth=Class, estimate=.pred_class)
conf_mat_rf<-predict_at_threshold(rf_final_model, Testdat, rf_best_thresh)%>%
  conf_mat(truth=Class, estimate=.pred_class)
conf_mat_svm<-predict_at_threshold(svm_final_model, Testdat, svm_best_thresh)%>%
  conf_mat(truth=Class, estimate=.pred_class)

#LDA
cmlda<-autoplot(conf_mat_lda, type="heatmap") + labs(title="LDA Matrix")

#QDA
cmqda<-autoplot(conf_mat_qda, type="heatmap") + labs(title="QDA Matrix")

#KNN
cmk<-autoplot(conf_mat_knn, type="heatmap") + labs(title="KNN Matrix")

#LR
cmlr<-autoplot(conf_mat_logreg, type="heatmap") + labs(title="Logreg Matrix")

#RF
cmrf <- autoplot(conf_mat_rf, type="heatmap") + 
  labs(title="Random Forest Matrix")

#SVM
cmsvm <- autoplot(conf_mat_svm, type="heatmap") + 
  labs(title="SVM Poly Matrix")

(cmlda+cmqda)/
(cmk+cmlr)/
(cmrf+cmsvm) 
```


```{r}
compare_test_results <- function(model, new_data, truth_col, model_name = "Model") {
  preds <- predict(model, new_data, type = "prob") %>%
    bind_cols(predict(model, new_data)) %>%
    bind_cols(new_data %>% select({{truth_col}})) %>%
    rename(truth = {{truth_col}})  # standard name for yardstick

  metrics_tbl <- metric_set(precision, recall, f_meas, roc_auc)(
    data = preds,
    truth = truth,
    estimate = .pred_class,
    .pred_Blue_Tarp)

  result <- metrics_tbl %>%
    select(.metric, .estimate) %>%
    pivot_wider(names_from = .metric, values_from = .estimate) %>%
    mutate(Model = model_name) %>%
    select(Model, precision, recall, f_meas, roc_auc)
  
  return(result)}
```


```{r}
#calc results 
#LDA here
lda_lin_results <- compare_test_results(lda_final_model, Testdat, truth_col = Class, model_name = "LDA")
#QDA here 
qda_lin_results <- compare_test_results(qda_final_model, Testdat, truth_col = Class, model_name = "QDA")
knn_results <- compare_test_results(knn_final_model, Testdat, truth_col = Class, model_name = "KNN")
lr_results <- compare_test_results(logreg_final_model, Testdat, truth_col = Class, model_name = "Log Reg")
#RF here 
randfor_results <- compare_test_results(rf_final_model, 
                                   Testdat, 
                                   truth_col=Class,
                                   model_name="Random Forest")
#SVM here
svm_R_results <- compare_test_results(svm_final_model,
                                    Testdat,
                                    truth_col=Class,
                                    model_name="SVM Poly")

all_results <- bind_rows(lda_lin_results, qda_lin_results, knn_results, lr_results, 
                         randfor_results, svm_R_results)
all_results%>%
  as_tibble() %>%
  kable(
    caption = "Comparison of All Models",
    digits  = 3
  )  %>% kableExtra::kable_styling(full_width=FALSE) 
```

